---
title: "Group Project"
output:
  html_document:
    df_print: paged
---

```{r, include=F, echo=F}
# Clear workspace
rm(list = ls()) 

# your code will NOT be included in the html for this assignment
knitr::opts_chunk$set(echo=FALSE) 
# your output will NOT be included in the html for this assignment (except figures at the end)
knitr::opts_chunk$set(include=FALSE) 

# load packages we need for this lab
library(mosaic, warn.conflicts = FALSE) 
library(ggformula, warn.conflicts = FALSE)
library(Lock5Data, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)


# Load your data
#CrimeDataRaw <- read.csv(file.choose())
CrimeDataRaw <- read.csv("sp24DenverCrimeData.csv")

# Perform any necessary transformations

```

## Introduction

For this project we have considered a sample of 75,000 individual data points that have been collected from the City of Denver's crime database, dating back for the previous 5 calendar years. Each individual data point includes data such as neighborhood insights of where the crime was committed, such as median income, demographics, etc, and details of the crime itself, such as the description of the crime, time of the crime, etc. We are curious about whether there is a positive association between the percentage of residents that are impoverished and whether residents are in school. Additionally, if the number of crimes are equally spread across the three crime types (non-violent, violent, and property).

## Is there a positive association between the percent of a neighborhood in poverty and the percent of the population in school?

This question did not require data manipulation. We have 75,000 observations available for this question, however each observation is representative of the percent of its neighborhood's population in school and in poverty so many of the entries in the data set are repeated.

Our hypotheses are:

$$H_0: \text{There is no association between the percent of a neighborhood in poverty and the percent of the population in school).}$$ 
$$H_a: \text{There is a positive association between the percent of a neighborhood in poverty and the percent of the population in school).}$$

We created a scatter plot to visualize these two variables. The majority of the cases had a poverty percentage of between 0% and 10% and a percentage of school between 12.5% and 25%. It appears that there is a positive association between the two variables.

To test the hypothesis we performed an inference for correlation. The resulting R\^2 value was 0.02201. The percentage in poverty accounts for 2.201% of the variability in the percent of the population in school. There is no evidence that there is a positive association between the percent of a neighborhood in poverty and the percent of the population in school.

## Are the different types of crime (non-violent, violent, and property) not equally likely to occur.

We have 75,000 observations available for this question.

Data manipulation: We began by transforming the variable 'crime_category' into three, which we have created a new variable called 'categorized_crime.' Data points that received the 'violent' type for 'crime_category,' includes crimes such as assault, robbery, homicide, etc, with data points receiving 'property_crimes' are theft, arson, etc. The rest are categorized a 'non-violent.' After. performing those operations in creating a new variable, we are left with being able to answer the following hypothesis:

$$H_0: \text{The distribution of crimes is equal across the three crime categories (non-violent, violent, and property).}$$ 
$$H_a: \text{The distribution of crimes is not equal across the three crime categories (non-violent, violent, and property).}$$

We created a bar chart representing the three types of crimes in our new 'crime_categorized' variable. Across all of the crime groups, we can easily notice that all of the crimes are not equal in proportion to each other.

To test our hypothesis, we have performed a chi-squared test of independence due to having three levels in our 'crime_categorized' variable. The p-value that is of result is 0. We have strong evidence to reject our null, and that the distribution of crimes is not equal across the three crime categories (non-violent, violent, and property).

\newpage

## Figures

```{r, include = T}
ResearchQuestion1 <- CrimeDataRaw
gf_point(percent_in_school ~ percent_poverty, data = ResearchQuestion1)

```

```{r, include = T}

violent_crimes <- c("assault-simple", "weapon-unlawful-discharge-of", "menacing-felony-w-weap", "agg-aslt-shoot",
                    "robbery-bank", "robbery-street", "robbery-business", "robbery-car-jacking", "robbery-residence",
                    "h omicide-other", "aggravated-assault", "kidnap-adult-victim", "weapon-other-viol", "agg-aslt-police-weapon",
                    "false-imprisonment", "weapon-fire-into-occ-bldg", "weapon-fire-into-occ-veh",
                    "explosive-incendiary-dev-use", "explosive-incendiary-dev-pos", "threats-to-injure", "assault-police-simple",
                    "homicide-family")

property_crimes <- c("theft-parts-from-vehicle", "theft-items-from-vehicle", "burglary-residence-by-force",
                     "burglary-residence-no-force", "burglary-business-no-force", "burglary-business-by-force",
                     "theft-other", "theft-bicycle", "theft-from-bldg", "theft-shoplift", "vehicle-theft", "theft-of-motor-vehicle",
                     "burglary-safe", "theft-embezzle", "burglary-vending-machine", "burg-auto-theft-busn-no-force",
                     "burg-auto-theft-resd-no-force", "burg-auto-theft-busn-w-force", "burg-auto-theft-resd-w-force",
                     "theft-purse-snatch-no-force", "theft-stln-vehicle-trailer", "theft-fail-return-rent-veh",
                     "theft-unauth-use-of-ftd", "theft-of-rental-property", "theft-pick-pocket", "arson-other",
                     "arson-business", "arson-residence", "arson-vehicle", "arson-public-building", "theft-of-services",
                     "theft-gas-drive-off", "burglary-poss-of-tools", "theft-from-mails", "stolen-property-buy-sell-rec",
                     "property-crimes-other")

# Function to categorize crimes
categorize_crime <- function(crime_type) {
  if (crime_type %in% violent_crimes) {
    return("Violent")
  } else if (crime_type %in% property_crimes) {
    return("Property")
  } else {
    return("Nonviolent")
  }
}

# Copy the data frame to a new variable.
ResearchQuestion2 <- CrimeDataRaw

# Applying the categorization to the copy
ResearchQuestion2$crime_category <- sapply(ResearchQuestion2$offense_type_id, categorize_crime)

# Creating the bar plot
ggplot(ResearchQuestion2, aes(x = crime_category, fill = crime_category)) +
  geom_bar() +  # This geom_bar plots a bar for each category of crime
  labs(title = "Distribution of Crime Categories",
       x = "Crime Category",
       y = "Frequency",
       fill = "Category") +
  theme_minimal() +  # Using a minimal theme for better aesthetics
  theme(plot.title = element_text(hjust = 0.5))  # Centering the plot title


```

# Appendix

*Put your R code to perform the hypothesis tests here. It must appear in your report with the output*

```{r, include = T, echo = T}

summary(lm(percent_in_school ~ percent_poverty, data=ResearchQuestion1))


```

```{r, include = T, echo = T}
# calculate observed stats
observed <- table(ResearchQuestion2$crime_category)
# null proportion, all types of crime are equal
# could be swapred for national data to see if denver matches the national expected values
expected_prob <- c(1/3, 1/3, 1/3)

# chi squared inference
xchisq.test(x = observed, p = expected_prob)
```
