---
title: "Group 4 Final Project"
output:
  html_document:
    df_print: paged
---

```{r, include=F, echo=F}
# Clear workspace
rm(list = ls()) 

# your code will NOT be included in the html for this assignment
knitr::opts_chunk$set(echo=FALSE) 
# your output will NOT be included in the html for this assignment (except figures at the end)
knitr::opts_chunk$set(include=FALSE) 

# load packages we need for this lab
library(mosaic, warn.conflicts = FALSE) 
library(ggformula, warn.conflicts = FALSE)
library(Lock5Data, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)


# Load your data
#CrimeDataRaw <- read.csv(file.choose())
CrimeDataRaw <- read.csv("sp24DenverCrimeData.csv")

# Perform any necessary transformations

```

## Introduction

For this project, we have analyzed a sample of 75,000 individual data points sourced from the City of Denver's crime database, covering the preceding 5 calendar years. Each data point encompasses neighborhood attributes where the crime occurred, such as median income, demographics, and more, along with specifics of the crime itself, including its description and time. We aim to investigate whether there exists a positive correlation between the percentage of impoverished residents and their school attendance. Furthermore, we seek to determine if the occurrence of crimes is evenly distributed among the three crime categories: non-violent, violent, and property.

## Is there an association between the percent of a neighborhood in poverty and the percent of the population in school?

This question did not require data manipulation. We have 75,000 observations available for this question, however each observation is representative of the percent of its neighborhood's population in school and in poverty so many of the entries in the data set are repeated.

Our hypotheses are:

$$H_0: \rho =0 \text{ (No association between the percent in poverty and the percent of the population in school).}$$ 
$$H_a: \rho \neq 0 \text{ (There is an association between the percent in poverty and the percent of the population in school).}$$

We created a scatter plot to visualize these two variables. The majority of the cases had a poverty percentage of between 0% and 10% and a percentage of school between 12.5% and 25%. It appears that there is a positive association between the two variables.

To test the hypothesis we performed a correlation test. The resulting $\rho$ value was 0.148364, and the resulting p-value was 0. There is strong evidence using a significance level of 0.05 that there is an association between the percent of a neighborhood in poverty and the percent of the population in school. However, because $\rho=0.148364$ the association between the percent in poverty and the percent of the population in school is weak. 

## Are the different types of crime (non-violent, violent, and property) not equally likely to occur.

We have 75,000 observations available for this question.

Data manipulation: We began by categorizing the variable 'crime_category' into three groups from which we created a new variable called 'categorized_crime.' Data points that received the 'violent' type for 'crime_category' includes crimes such as assault, robbery, homicide, with data points receiving 'property_crimes' are theft, arson, etc. The rest are categorized a 'non-violent.' After performing those operations, we are left with being able to answer the following hypothesis:

$$H_0: \text{The distribution of crimes is equal across the three crime categories (non-violent, violent, and property).}$$ 
$$H_a: \text{The distribution of crimes is not equal across the three crime categories (non-violent, violent, and property).}$$

We created a bar chart representing the three types of crimes in our new 'crime_categorized' variable. Across all of the crime groups, we can easily notice that all of the crimes are not equal in proportion to each other.

To test our hypothesis, we have performed a chi-squared test of independence due to having three levels in our 'crime_categorized' variable. The p-value that is of result is 0. We have strong evidence to reject our null, and that the distribution of crimes is not equal across the three crime categories (non-violent, violent, and property).

\newpage

## Figures

```{r, include = T}
ResearchQuestion1 <- CrimeDataRaw
ggplot(ResearchQuestion1, aes(x = percent_poverty, y = percent_in_school)) +
  geom_point() +  # Adding points
  labs(title = "Relationship Between Poverty and School Attendance",
       x = "Percent Poverty",  # X-axis label
       y = "Percent in School")  # Y-axis label

```

```{r, include = T}

violent_crimes <- c("assault-simple", "weapon-unlawful-discharge-of", "menacing-felony-w-weap", "agg-aslt-shoot",
                    "robbery-bank", "robbery-street", "robbery-business", "robbery-car-jacking", "robbery-residence",
                    "h omicide-other", "aggravated-assault", "kidnap-adult-victim", "weapon-other-viol", "agg-aslt-police-weapon",
                    "false-imprisonment", "weapon-fire-into-occ-bldg", "weapon-fire-into-occ-veh",
                    "explosive-incendiary-dev-use", "explosive-incendiary-dev-pos", "threats-to-injure", "assault-police-simple",
                    "homicide-family")

property_crimes <- c("theft-parts-from-vehicle", "theft-items-from-vehicle", "burglary-residence-by-force",
                     "burglary-residence-no-force", "burglary-business-no-force", "burglary-business-by-force",
                     "theft-other", "theft-bicycle", "theft-from-bldg", "theft-shoplift", "vehicle-theft", "theft-of-motor-vehicle",
                     "burglary-safe", "theft-embezzle", "burglary-vending-machine", "burg-auto-theft-busn-no-force",
                     "burg-auto-theft-resd-no-force", "burg-auto-theft-busn-w-force", "burg-auto-theft-resd-w-force",
                     "theft-purse-snatch-no-force", "theft-stln-vehicle-trailer", "theft-fail-return-rent-veh",
                     "theft-unauth-use-of-ftd", "theft-of-rental-property", "theft-pick-pocket", "arson-other",
                     "arson-business", "arson-residence", "arson-vehicle", "arson-public-building", "theft-of-services",
                     "theft-gas-drive-off", "burglary-poss-of-tools", "theft-from-mails", "stolen-property-buy-sell-rec",
                     "property-crimes-other")

# Function to categorize crimes
categorize_crime <- function(crime_type) {
  if (crime_type %in% violent_crimes) {
    return("Violent")
  } else if (crime_type %in% property_crimes) {
    return("Property")
  } else {
    return("Nonviolent")
  }
}

# Copy the data frame to a new variable.
ResearchQuestion2 <- CrimeDataRaw

# Applying the categorization to the copy
ResearchQuestion2$crime_category <- sapply(ResearchQuestion2$offense_type_id, categorize_crime)

# Creating the bar plot
ggplot(ResearchQuestion2, aes(x = crime_category, fill = crime_category)) +
  geom_bar() +  # This geom_bar plots a bar for each category of crime
  labs(title = "Distribution of Crime Categories",
       x = "Crime Category",
       y = "Frequency",
       fill = "Category") +
  theme_minimal() +  # Using a minimal theme for better aesthetics
  theme(plot.title = element_text(hjust = 0.5))  # Centering the plot title


```

# Appendix

```{r, include = T, echo = T}

cor.test(percent_in_school ~ percent_poverty, data=ResearchQuestion1)


```

```{r, include = T, echo = T}
# calculate observed stats
observed <- table(ResearchQuestion2$crime_category)
# null proportion, all types of crime are equal
# could be swapred for national data to see if denver matches the national expected values
expected_prob <- c(1/3, 1/3, 1/3)

# chi squared inference
xchisq.test(x = observed, p = expected_prob)
```
