---
title: "Group Project"
output:
  html_document:
    df_print: paged
---

```{r}
# Clear workspace
rm(list = ls()) 

# your code will NOT be included in the html for this assignment
knitr::opts_chunk$set(echo=FALSE) 
# your output will NOT be included in the html for this assignment (except figures at the end)
knitr::opts_chunk$set(include=FALSE) 

# load packages we need for this lab
library(mosaic, warn.conflicts = FALSE) 
library(ggformula, warn.conflicts = FALSE)
library(Lock5Data, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)


# Load your data
CrimeDataRaw <- read.csv(file.choose())

# Perform any necessary transformations


```


## Introduction

*Introduce the dataset in your own words.  Discuss, briefly, the motivation and curiousity behind your choice of research questions.  *  
  
## Research Question 1

*Write the research question in the section title and clearly detail any data manipulations you performed to be able to answer this research question. Do not include R code; use complete sentences to describe what you did. * 
  
   - *Write the relevant hypothesis test and clearly define the parameter of interest in the context of your research question.* 
   - *Discuss what kind of data visualization you created for this question, and what you learned from it as it relates to the research question.  You will put this visualization at the end of the document.*
  - *Introduce the inference procedure and report the p-value, as well as your conclusion in context.  Make sure you discuss the strength of the evidence.*
  
## Research Question 2

*Write the research question in the section title and clearly detail any data manipulations you performed to be able to answer this research question. Do not include R code; use complete sentences to describe what you did. * 
  
   - *Write the relevant hypothesis test and clearly define the parameter of interest in the context of your research question.* 
   - *Discuss what kind of data visualization you created for this question, and what you learned from it as it relates to the research question.  You will put this visualization at the end of the document.*
  - *Introduce the inference procedure and report the p-value, as well as your conclusion in context.  Make sure you discuss the strength of the evidence.*
  
\newpage

## Figures

```{r, include = T}
ResearchQuestion1 <- CrimeDataRaw
gf_point(percent_in_school ~ percent_poverty, data = ResearchQuestion1)

```

```{r, include = T}

violent_crimes <- c("assault-simple", "weapon-unlawful-discharge-of", "menacing-felony-w-weap", "agg-aslt-shoot",
                    "robbery-bank", "robbery-street", "robbery-business", "robbery-car-jacking", "robbery-residence",
                    "homicide-other", "aggravated-assault", "kidnap-adult-victim", "weapon-other-viol", "agg-aslt-police-weapon",
                    "false-imprisonment", "weapon-fire-into-occ-bldg", "weapon-fire-into-occ-veh",
                    "explosive-incendiary-dev-use", "explosive-incendiary-dev-pos", "threats-to-injure", "assault-police-simple",
                    "homicide-family")

property_crimes <- c("theft-parts-from-vehicle", "theft-items-from-vehicle", "burglary-residence-by-force",
                     "burglary-residence-no-force", "burglary-business-no-force", "burglary-business-by-force",
                     "theft-other", "theft-bicycle", "theft-from-bldg", "theft-shoplift", "vehicle-theft", "theft-of-motor-vehicle",
                     "burglary-safe", "theft-embezzle", "burglary-vending-machine", "burg-auto-theft-busn-no-force",
                     "burg-auto-theft-resd-no-force", "burg-auto-theft-busn-w-force", "burg-auto-theft-resd-w-force",
                     "theft-purse-snatch-no-force", "theft-stln-vehicle-trailer", "theft-fail-return-rent-veh",
                     "theft-unauth-use-of-ftd", "theft-of-rental-property", "theft-pick-pocket", "arson-other",
                     "arson-business", "arson-residence", "arson-vehicle", "arson-public-building", "theft-of-services",
                     "theft-gas-drive-off", "burglary-poss-of-tools", "theft-from-mails", "stolen-property-buy-sell-rec",
                     "property-crimes-other")

# Function to categorize crimes
categorize_crime <- function(crime_type) {
  if (crime_type %in% violent_crimes) {
    return("Violent")
  } else if (crime_type %in% property_crimes) {
    return("Property")
  } else {
    return("Nonviolent")
  }
}

# Copy the data frame to a new variable.
ResearchQuestion2 <- CrimeDataRaw

# Applying the categorization to the copy
ResearchQuestion2$crime_category <- sapply(ResearchQuestion1$offense_type_id, categorize_crime)

# Creating the bar plot
ggplot(ResearchQuestion2, aes(x = crime_category, fill = crime_category)) +
  geom_bar() +  # This geom_bar plots a bar for each category of crime
  labs(title = "Distribution of Crime Categories",
       x = "Crime Category",
       y = "Frequency",
       fill = "Category") +
  theme_minimal() +  # Using a minimal theme for better aesthetics
  theme(plot.title = element_text(hjust = 0.5))  # Centering the plot title


```

# Appendix

*Put your R code to perform the hypothesis tests here.  It must appear in your report with the output*

```{r, include = T, echo = T}

ResearchQuestion1Model = lm(formula = percent_in_school ~ percent_poverty, data = ResearchQuestion1)
summary(ResearchQuestion1Model)


```

```{r, include = T, echo = T}
ResearchQuestion2Crimes <- table(ResearchQuestion2$crime_category)
Alternatives <- c(30000, 30000, 30000)
ResearchQuestion2Table <- rbind(ResearchQuestion2Crimes, Alternatives)
xchisq.test(ResearchQuestion2Table)
```




